---
title: "Survey Responses - Political & Environmental"
author: "Chase Tarr"
date: "11/29/2021"
output: html_document
---

### Introduction
This data summarizes the survey results from a surveyed population in which individuals self-reported their political preference (Conservative or Liberal), and if they identified themselves as an environmentalist by answering "Yes", "No", or "Unsure". We are seeking to determine if there is a significant association between an individuals political preference and whether or not they identify as an environmentalist. 
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)

library(tidyverse)
library(broom)
library(janitor)
library(here)
library(kableExtra)
library(DT)
```

```{r}
survey_data <- read_csv(here("data", "conservation_survey.csv")) %>% 
  clean_names() %>% 
   select(env_id, pol_or) %>% 
  mutate(Political_Party = case_when(
    pol_or %in% c("1","2","3") ~ "Conservative",
    pol_or %in% c("5", "6", "7") ~ "Liberal")) %>% 
  mutate(areyou_enviro = case_when(
    env_id == "1" ~ "Yes",
    env_id == "2" ~ "No",
    env_id == "3" ~ "Unsure")) %>% 
  filter(Political_Party %in% c("Conservative", "Liberal"))
```


```{r}
party_counts <- survey_data %>% 
  janitor::tabyl(Political_Party, areyou_enviro)

party_proportions <- party_counts %>% 
  adorn_percentages() %>% 
  adorn_pct_formatting(digits = 2) %>% 
  adorn_ns()

datatable(party_proportions,
      caption = "Table 1: Political party preference correlating to the proportion of proclaimed Environmentalists")
```

```{r}
party_cs <- party_counts %>% 
  column_to_rownames(var = "Political_Party")
```

```{r}
party_x2 <- chisq.test(party_cs)

partyx2_tidy <- tidy(party_x2)
```

We used Pearson's Chi-Squared Test to evaluate the relationship between political preference and whether or not a given person considers themselves as an environmentalist. Upon analyzing the data, it is evident that there is a significant association between political association (conservative or liberal leaning) and if a person considers themselves an environmentalist($\chi$\^2\^(`r partyx2_tidy$parameter`) = `r round(partyx2_tidy$statistic,2)`, *p* = `r partyx2_tidy$p.value`). Within the surveyed population, `r party_proportions$No [1]` of conservative identifying individuals *would not* label themselves as an environmentalist, whereas `r party_proportions$Yes [2]` of liberal identifying individuals *would* label themselves as environmentalist showing a clear correlation between the two.


**Data Citation:** Andrews Forest LTER Site and C. Batavia. 2019. Influences on charitable giving for conservation: Online survey data of 1,331 respondents across the US, August 2017 ver 3. Environmental Data Initiative. https://doi.org/10.6073/pasta/926e6270e324a1322a900da14d38b96c
