---
title: "Mono Lake In Comparison to Sea Level from 1850-2017"
author: "Chase Tarr"
date: "11/29/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
library(tidyverse)
library(broom)
library(janitor)
library(here)
library(kableExtra)
library(DT)
library(ggbeeswarm)
library(plotly)
library(effsize)
library(gganimate)
library(gapminder)
```

```{r}
mono_lake <- read_csv(here("data", "mono_lake_annual_levels.csv")) %>% 
  clean_names()

mono_lake <- mono_lake[!apply(is.na(mono_lake) | mono_lake == "", 1, all),]

mono_lake_2 <- mono_lake %>% slice(-c(1,2,3,4)) %>% 
  rename(lake_level_ft = x2,
         Year = mono_lake_level_1850_present) %>% 
  select(Year, lake_level_ft)

mono_lake_2$lake_level_ft <- as.numeric(as.character(mono_lake_2$lake_level_ft))
mono_lake_2$Year <- as.numeric(as.character(mono_lake_2$Year))
```

```{r}
ml <- ggplot(data = mono_lake_2, aes(x = Year, y = lake_level_ft)) +
  geom_point(aes(color = Year)) +
  theme(axis.text.x = element_text(angle = 90, hjust = 1)) + 
  scale_x_continuous(breaks = seq(1850, 2017, by = 15)) +
  labs(title = "Mono Lake Level Per Year (ft above sea level)", 
       x = "Year",
       y = "Level (ft)",
       caption = "Figure 1: Source: https://www.monobasinresearch.org/") +
  theme(plot.title = element_text(size = 15,
                                  face = "bold",
                                  family = "American Typewriter",
                                  color = "deepskyblue3",
                                  hjust = 0.5,
                                  lineheight = 1.2),
        plot.subtitle = element_text(size = 10,
                                     family = "American Typewriter",
                                     face = "bold",
                                     hjust = 0.5),
        plot.caption = element_text(size = 8,
                                    hjust = 1),
        legend.position = c(0.93,0.7),
        legend.background = element_blank())
```

```{r}
point1 <- mono_lake_2 %>% 
  filter(lake_level_ft == "6376")

point2 <- mono_lake_2 %>% 
  filter(Year == "1918")

line1 <- mono_lake_2 %>% 
  filter(Year %in% c(1918:1977))

point3 <- mono_lake_2 %>% 
  filter(Year == "1941")
```

```{r}
ml +
  geom_point(data = point1,
             aes(x = Year, y = lake_level_ft, color = Year),
             color = "red",
             size = 5) +
  geom_point(data = point2,
             aes(x = Year, y = lake_level_ft, color = Year),
             color = "limegreen",
             size = 5) +
  geom_point(data = point3,
             aes(x = Year, y = lake_level_ft, color = Year),
             color = "darkorange3",
             size = 5) +
  geom_line(data = line1, color = "red", alpha = 0.5) +
  geom_label(label = "Highest Level",
             x = 1899, y = 6427,
             label.padding = unit(0.12, "lines"),
             label.size = 0.35,
             color = "black",
             fill = "limegreen") +
  geom_label(label = "Land Bridge Emerges",
             x = 1951, y = 6378,
             label.padding = unit(0.12, "lines"),
             label.size = 0.35,
             color = "black",
             fill = "red") +
  geom_label(label = "LA Extends Aqueduct",
             x = 1961, y = 6421,
             label.padding = unit(0.12, "lines"),
             label.size = 0.15,
             color = "black",
             fill = "darkorange3") +
  geom_hline(yintercept = 6376, linetype = "dashed", color = "red")
```

**Figure 1** represents data points of the recorded feet above sea level for Mono Lake each year.The *highest recorded value* of 6427 feet above sea level is represented by a *green dot*. This value is then traced with a red line to the point at which a *land bridge emerges* represented by the *red dot in 1977* and shown across time by the red dashed line. This land bridge connects the mainland to the islands which allows predators to access major bird nesting sites negatively affecting migratory bird species and disrupting local ecosystems. The decrease from 1918 to 1977 can be attributed to human interaction in which the city of Los Angeles re-routed a portion of the water for urban purposes that fed the Mono County Basin prior. This, compounded with recent drought, has caused the decline seen in the last century. The initial decline just after 1918 was due to the initial "draining" as some call it in which Los Angeles first began funneling water into the city. The lake was first recorded at 6405 feet above sea level in 1850, and currently sits near 6381 above sea level which was the last recording in 2017, but has recently reached as low as 6372.30 feet above sea level in the 80's. A *small increase* just before 1940 was quickly negated by a sharp decline from 1941 to 1982 due to the Los Angeles department of Water and Power extending the Los Angeles Aqueduct System more north (shown by the *orange dot*) into the Mono Basin therefore pulling more water that would otherwise go into the lake.

**Data Citation:** “Mono Basin Clearinghouse: Building a Digital Library for Better Resource Management.” Mono Basin Clearinghouse: Building a Digital Library for Better Resource Management, https://www.monobasinresearch.org/. 
